Intersection of non-matching grids of different dimensions

### 2016-09-25 paper_intersections.tex

\section{Element Intersections}
\label{sec:elements_intersections}
In this section, we present algorithms for computing intersection of a pair of simplicial elements of a~different dimension in the 3D ambient space. 
In particular we are interested in the intersections in 1D-2D, 1D-3D, 2D-2D, 2D-3D cases.
It will be suitable to denote $S_i$ a simplicial element with $i$ vertices (of dimension $i-1$).
In general, an intersection can be a point, a line segment or a polygon.
All these three objects are represented as a vector of vertices called \emph{intersection points} 
(later denoted IP, or IPs in plural).

The fundamental idea is to compute intersection of 1D-2D simplices using the 
\plucker coordinates and reduce all other cases to this one.

@@

Next, we shall discuss the case when some of the permuted inner products are zero. In fact, we test if any barycentric coordinate is under given tolerance 
($\abs{w_i} \le \epsilon$). Various \emph{pathological} cases are detected and further topological 
information is atached to the intersection data.  In the case of one zero coordinate (non-coplanar intersection on the edge)
or two zero coordinates (intersection in the vertex), we add the local edge index or the local vertex index to the intersection, respectively.
When all coordinates are (close to) zero, the line and the triangle are (nearly) coplanar. For every edge $s_i$ we compute
its intersection with the line segment $p$ in terms of the local coordinates $t_i$ and $t_p$. Only intersections with $t_i \in (-\epsilon, 1+\epsilon)$
are accepted, for $t_i$ close to $0$ and $1$ we append the vertex index to the intersection. Edges parallel to $p$ or lying on the $p$ are skipped.

Finally, we obtain zero, one, or two intersection points with additional topological information.
The intersection point data structure keeps for each of the two intersecting elements:
\begin{itemize}
 \item Barycentric coordinate of the intersection.
 \item Dimension of the most specific subelement (face, edge, vertex) where the point lies.
 \item Local index of the most specific subelement.
\end{itemize}
Further, we store the sign of the permuted inner product in order to detect various special and degenerate cases higher dimensional cases.

@@

\begin{algorithm}
  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{input data}
  \KwOut{List of IPs sorted output data}  
  $I =\emptyset$\tcp*{Intersection polygon.}
  $F(:,:)=-1$\tcp*{Unset entries of the face connection table.}
  \For{side $s$ of $S_3$}{ 
    \nllabel{line:1_3}
    $L$ = intersection($s, S_4$)  \tcp*{List of max.\ two points.}
    process point($L[0]$, 0)\; 
    process point($L[-1]$, 1)
    \notePE{Isn't it clearer to write the two process point lines into cycle over L?}
  }
  \lIf{ $I[0] == I[-1]$ }{remove $I[-1]$}
  \notePE{I am not sure if I understand the notation '($[-1]$)', can be explained?
The last line then checks if first and last IPs, if they are not duplicates..??}
  \caption{2d-3d intersection, points on triangle boundary}
  \label{algo:colect_23_ip_triangle}
\end{algorithm}

\begin{algorithm}
  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{intersection point $p$, face relation indicator $i$}
  \KwOut{update of $I$ and $F$}
  \lIf{ $p \ne I[-1]$ }{append $p$ to $I$}
  \notePE{I am missing the check whether IP is vertex of $S_3$ as referred in text.
  So what if it is a vertex?}\;
  set $j$ to index of $p$ in $I$\;
  \lIf{ $p$ is on face $f$}{ F[i,f]=j}
  \uElseIf{ $p$ is on edge $e$}{
    compute sign of intersection($e$, $S_3$)\;
    sort faces $f_0$, $f_1$ coincident with $e$\;
    F[i, $f_i$]=j
  }
  \ElseIf{ $p$ is on vertex $v$ of $S_4$}{
      \For{face $f$ coincident with $v$}{
         \lIf{ $F[i,f]==-1$}{F[i,f]=j}
      }
  }
  mark all edges of $S_4$ the $p$ lies on

  \caption{process point( $p$, $i$ )}
  \label{algo:process_point}
\end{algorithm}

\begin{algorithm}
  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{input data}
  \KwOut{List of IPs sorted output data}  
  \For{unmarked edge $e$ of $S_4$}{
    $p$ = intersection($e, S_3$)\;
    append $p$ to $I$\;
    set $j$ to index of $p$ in $I$\;
    \uIf{ $p$ is inside of $e$}{ 
        sort faces $f_0$, $f_1$ coincident with $e$\;
        F[0, $f_0$]=j\;
        F[1, $f_1$]=j
    } %else
    \Else($p$ on the vertex $v$ of $S_4$){       
      \For{face $f$ coincident with $v$}{
         \lIf{ $F[i,f]==-1$}{F[i,f]=j}
      }
      mark all edges of $S_4$ coincident with $v$
    }
  }
  \caption{2d-3d intersection, points in triangle interior}
  \label{algo:collect_23_ip_edges}
\end{algorithm}

\begin{algorithm}
  \SetKw{To}{To}
  \DontPrintSemicolon
  \KwIn{all points in $I$}
  \KwOut{polygon in $J$ in correct order \notePE{J should be K ?}}  
  $P=(1,\dots, n-1,0)$\tcp{array of sucessors}
  \For{ $f$ is face of $S_4$}{
    \lIf{$F[1,f] \ne F[1,f]$}{$P[ F[1,f] ] = F[0, f]$} 
    \notePE{condition should be $F[0,f] \ne F[1,f]$ ? I don't understand: successor indices should be switched?
    This way there would be same two indices..}
  }
  $i=0$

  \lFor{$n=0$ \To $\abs{I}-1$}{
    $K[n]=I[i]$
    $i=P[i]$
  }    
  \caption{2d-3d intersection, finish sort of points}
  \label{algo:order_23}
\end{algorithm}

@@

\subsection{Tracking boundary of the intersection polygon}

@@

\notePE{We need proper definitions of terms we use: \\
% \emph{component} is a set of connected elements of the same lower dimension (1D, 2D)\\
% \emph{bulk} is a set of 3D elements\\
% or slave, master...
\emph{candidates pair}  is a pair of a component element and a bulk element, that might intersect
each other (due to intersection of their bounding boxes or prolongation result)
\emph{pathologic, special, degenerate case ??}
}

@@

\subsection{Prolongation algorithm}
Consider now a complex mesh of combined dimensions consisting of \emph{components}, which are sets of connected
elements of the same lower dimension (1D, 2D), in the space of 3D elements, which we shall call a \emph{bulk}. 
Obtaining all of the 1D-3D and 2D-3D intersections is based on finding the first two elements intersecting each other.
Then we can prolongate the intersection by investigating neighbouring elements.
%
\begin{figure}[!htb]
%   \vspace{0pt}
  \centering    
    \includegraphics[width=\textwidth]{\figpath prolongation_scheme.pdf}
  \caption{Prolongation algorithm for 1D-2D and 2D-3D intersections. }
  \label{fig:prolongation}
\end{figure}

>>>>

\section{Element Intersections}
\label{sec:elements_intersections}
In this section, we present algorithms for computing intersection of a pair of simplicial elements of a~different dimension in the 3D ambient space.
In particular we are interested in intersection of 1D-2D, 1D-3D, 2D-2D, 2D-3D simplices. In general the intersection can be a point, a line segment or a polygon.
All these three objects are represented as a vector of vertices called \emph{intersection points}. The fundamental idea is to compute intersection of 1D-2D simplices using the 
\plucker coordinates and reduce all other cases to this one.

The intersection point data structure keeps for each of the two intersection simplices:
\notePE{Subsimplex or subelement?}
\begin{itemize}
 \item Barycentric coordinate of the intersection.
 \item Dimension of the most specific subsimplex (face, edge, vertex) where the point lies.
 \item Index of the most specific subelement.
\end{itemize}
Further, we store the sign of the permuted inner product
in order to detect various special and degenerate cases.

@@

Next, we shall discuss the case when some of the permuted inner products are zero. In fact, we test if any barycentric coordinate is under given tolerance 
($\abs{w_i} \le \epsilon$). Various \emph{pathological} cases are detected and further topological 
information is atached to the intersection data.  In the case of one zero coordinate (non-coplanar intersection on the edge)
or two zero coordinates (intersection in the vertex), we add the local edge index or the local vertex index to the intersection, respectively.
When all coordinates are (close to) zero, the line and the triangle are (nearly) coplanar. For every edge $s_i$ we compute
its intersection with the line segment $p$ in terms of the local coordinates $t_i$ and $t_p$. Only intersections with $t_i \in (-\epsilon, 1+\epsilon)$
are accepted, for $t_i$ close to $0$ and $1$ we append the vertex index to the intersection. Edges parallel to $p$ or lying on the $p$ are skipped.
Finally, we obtain zero, one, or two intersection points with additional topological information.

@@

\begin{algorithm}
  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{input data}
  \KwOut{List of IPs sorted output data}  
  $I =\emptyset$\tcp*{Intersection polygon.}
  $F(:,:)=-1$\tcp*{Unset entries of the face connection table.}
  $V =\emptyset$\tcp*{$S_4$ vertex intersections.}
  \For{side $s$ of $S_3$}{ 
    \nllabel{line:1_3}
    $L$ = intersection($s, S_4$)  \tcp*{List of max.\ two points.}
    process point($L[0]$, 0)\; 
    process point($L[-1]$, 1)
    \notePE{Isn't it clearer to write the two process point lines into cycle over L?}
  }
  \lIf{ $I[0] == I[-1]$ }{remove $I[-1]$}
  \notePE{I am not sure if I understand the notation '($[-1]$)', can be explained?
The last line then checks if first and last IPs, if they are not duplicates..??}
  \caption{2d-3d intersection, points on triangle boundary}
  \label{algo:colect_23_ip_triangle}
\end{algorithm}

\begin{algorithm}
  \caption{process point( $p$, $i$ )}

  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{intersection point $p$, face relation indicator $i$}
  \KwOut{update of $I$ and $F$}
  \lIf{ $p \ne I[-1]$ }{append $p$ to $I$}
  \If{ $p$ is not vertex of $S_3$}{    
  \notePE{I am missing the check whether IP is vertex of $S_3$ as referred in text.
  So what if it is a vertex?}\;
    set $j$ to index of $p$ in $I$\;
    \lIf{ $p$ is on face $f$}{ F[i,f]=j}
    \uElseIf{ $p$ is on edge $e$}{
        F[i, $e$]=j
    }
    \ElseIf{ $p$ is on vertex $v$ of $S_4$}{
        F[i, $v$] =j
   }
  }    
  \label{algo:process_point}
\end{algorithm}

\begin{algorithm}
  \label{algo:collect_23_ip_edges}
  \caption{2d-3d intersection, points in triangle interior}

  \SetKw{And}{And}
  \DontPrintSemicolon
  \KwIn{input data}
  \KwOut{List of IPs sorted output data}  
  \For{edge $e$ of $S_4$}{
    $p$ = intersection($e, S_3$)\;
    \uIf{ $p$ is on boundary of $S_3$}{

    }
    \Else{
        append $p$ to $I$\;
        set $j$ to index of $p$ in $I$\;
        \uIf{ $p$ is inside of $e$}{ 
            sort faces $f_0$, $f_1$ coincident with $e$\;
            F[0, $f_0$]=j\;
            F[1, $f_1$]=j
        } %else
        \Else($p$ on the vertex $v$ of $S_4$){       
        append $j$ to $V$
        \For{face $f$ coincident with $v$}{
            \lIf{ $F[i,f]==-1$}{F[i,f]=j}
        }
        mark all edges of $S_4$ coincident with $v$
        }
     }   
  }
\end{algorithm}

\begin{algorithm}
  \caption{2d-3d intersection, finish sort of points}  
  \label{algo:order_23}

  \SetKw{To}{To}
  \DontPrintSemicolon
  \KwIn{all points in $I$}
  \KwOut{polygon in $K$ in correct order}  
  \If{\abs{V} = 3}{

  }
  $P=(1,\dots, n-1,0)$\tcp{array of sucessors}
  \For{ $f$ is face of $S_4$}{
    \lIf{$F[1,f] \ne F[1,f]$}{$P[ F[1,f] ] = F[0, f]$} 
    \notePE{condition should be $F[0,f] \ne F[1,f]$ ? I don't understand: successor indices should be switched?
    This way there would be same two indices..}
  }
  $i=0$

  \lFor{$n=0$ \To $\abs{I}-1$}{
    $K[n]=I[i]$
    $i=P[i]$
  }    
  \label{algo:order_23}
\end{algorithm}

@@

\subsection{Tracking boundary of the intersection polygon}
\notePE{I suggest to denote IP, IPs intersection point and its plural respectively.}

@@

\notePE{We need proper definitions of terms we use: \\
\emph{component} is a set of connected elements of the same lower dimension (1D, 2D)\\
\emph{bulk} is a set of 3D elements\\
% or slave, master...
\emph{candidates pair}  is a pair of a component element and a bulk element, that might intersect
each other (due to intersection of their bounding boxes or prolongation result)
\emph{pathologic, special, degenerate case ??}
}

@@

\subsection{Prolongation algorithm}
Consider now a complex mesh of combined dimensions consisting of \emph{components}, which are sets of connected
elements of the same lower dimension (1D, 2D), in the space of 3D elements. 
Obtaining all of the 1D-3D and 2D-3D intersections is based on finding the first two elements intersecting each other.
Then we can prolongate the intersection by investigating neighbouring elements.
%
\begin{figure}[!htb]
%   \vspace{0pt}
  \centering    
    \includegraphics[width=\textwidth]{\figpath prolongation_scheme.pdf}
  \caption{Prolongation algorithm for 1D-2D and 2D-3D intersections. }
  \label{fig:prolongation}
\end{figure}
